---
title: ""
author: ""
mail: "hello@datalobo.com"
twiter: "datalobo"
home: "www.datalobo.com"

output: 
  html_document:
    toc: true
    toc_float: true
    toc_smooth_scroll: true

css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("tidyverse","htmltools","reactable")
#Install packages not yet installed
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {install.packages(packages[!installed_packages])}

#Load packages
invisible(lapply(packages, library, character.only = TRUE))


blue_pal <- function(x) { rgb(colorRamp(c("#9fc7df", "#416ea4"))(x), maxColorValue = 255)}

make_color_pal <- function(colors, bias = 1) {
  get_color <- colorRamp(colors, bias = bias)
  function(x) rgb(get_color(x), maxColorValue = 255)
}

value_rating_color <- make_color_pal(c("#ffffff", "#f2fbd2", "#c9ecb4", "#93d3ab", "#35b0ab"), bias = 2)

load("/Users/danielgray/Desktop/projects/unions/union_table.RData")

table <- table %>% mutate(men_share = round(men_share * 100,0),
                         followers = ifelse(followers == 0,NA, followers)) %>%
                   mutate(followers = ifelse(`Trade union` == "RSPB Staff Association",NA,followers)) #These are the followers for the RSPB, not the union.

min_assets <- min(table$net_assets, na.rm = T)
max_assets <- ifelse(max(table$net_assets,na.rm = T) == "-Inf",0,max(table$net_assets,na.rm = T))

min_men <- min(table$men_share,na.rm =T)
max_men <- max(table$men_share,na.rm =T)

table <- table %>% mutate(
      normalised = (men_share - min_men) / (max_men - min_men),
      normalised = ifelse(is.na(normalised),0,normalised)) %>% ungroup() %>%
      mutate(score_colour = blue_pal(normalised)) %>%
      select(-c(normalised))

member_cols <- c("members", "men_share")
finance_cols <- c("net_assets", "income_statement")


```

## The UK's Trade Unions

### Introduction

The cost of living has increased by around 8% over the past year. To ward off a wage-price spiral, the government and the Bank of England have asked workers not to request big pay-rises.  Trade unions, meanwhile, have sought to negotiate pay-rises at least in line with inflation.   They want to protect their members’ standard of living and - as the recent strikes organised by RMT demonstrated - are willing to take industrial action.  

With the government pitched against unions some have drawn parallels with the 1970s, warning of a “summer of discontent”.  The comparison is exaggerated: unions are not as powerful as they were. But as disruption to public services is probable in the coming months, we wanted to profile the UK’s trade unions to help our readers interpret the inevitable news flow covering unions and potential industrial action. 

The interactive table below profiles the UK's trade unions. It includes data on how many members they have (and thus the scale of disruption they could potentially cause), how strong their finances are (and thus much firepower they have to spend on campaigns), and how many twitter followers they have (as a proxy for public interest in their cause).

There are some notable differences between the unions. Education unions have significantly more funds available to them than unions representing other sectors.  Nursing unions have  larger Twitter followings, and so should find it easier to generate interest in their cause (though RMT's humorous tweets in recent weeks have seen it climb the ranks). Unison and Unite have significantly more members than other unions. Whilst they are similar in that they both represent multiple different sectors, the gender split of their memberships differ significantly. Unison mostly represents public sector workers, which includes more women. 


### Data

The unions are sorted from most-members to fewest-members.  Click a column header to sort by an alternative metric.


```{r cars, echo = FALSE}

reactable(table,
      
          style = list( fontSize = "14px"),
          
columnGroups = list(
    colGroup(name = "Membership", columns = member_cols),
    colGroup(name = "Finances", columns = finance_cols)),

columns = list(
  
`Trade union` = colDef(
name = "Trade Union",
minWidth = 250
  
),

score_colour = colDef(show = F),

members = colDef(
  
  name = "Number",
  format = colFormat(separators = TRUE),
  maxWidth = 120,
  align = "center",
  
  cell = function(value) {
    
    if(is.na(value)) {""} else {
      scaled <- (log(value+1) - log(min(table$members,na.rm = T)+1)) / (log(max(table$members,na.rm = T)+1) - log(min(table$members,na.rm = T)))
      color <- value_rating_color(scaled)
      value <- format(round(value, 0),big.mark = ",", scientific = FALSE)
      div(class =  "spi-rating", style = list(background = color), value)}}
),

net_assets = colDef(
  
  name = "Net Assets",
  format = colFormat(separators = TRUE),
  maxWidth = 120,
  align = "center",
  
  cell = function(value) {
    
    if(is.na(value)) {""} else {
      scaled <- log(value+1-min_assets) / log(max_assets-min_assets + 1)
      color <- value_rating_color(scaled)
      value <- paste0("£",format(round(value, 0),big.mark = ",", scientific = FALSE))
      div(class =  "spi-rating", style = list(background = color), value)}}
),

income_statement = colDef(
  
  name = "Member Contributions",
  format = colFormat(separators = TRUE),
  maxWidth = 140,
  align = "center",
  
  cell = function(value) {
    
    if(is.na(value)) {""} else {
      scaled <- (log(value+1) - log(min(table$income_statement,na.rm = T)+1)) / (log(max(table$income_statement,na.rm = T)+1) - log(min(table$income_statement,na.rm = T)))
      color <- value_rating_color(scaled)
      value <- paste0("£",format(round(value, 0),big.mark = ",", scientific = FALSE))
      div(class =  "spi-rating", style = list(background = color), value)}}
),

followers = colDef(
  
  name = "Twitter Followers",
  format = colFormat(separators = TRUE),
  maxWidth = 120,
  align = "center",
  
  cell = function(value) {
    
    if(is.na(value)) {""} else {
      scaled <- log(value+1) / (log(max(table$followers,na.rm = T)+1))
      color <- value_rating_color(scaled)
      value <- format(round(value, 0),big.mark = ",", scientific = FALSE)
      div(class =  "spi-rating", style = list(background = color), value)}}
),


men_share = colDef(
                  
                  name = "% of which men",
                  maxWidth = 120,
                  align = "center",
                  
                  cell = JS("function(cellInfo) {
        const sliceColor = cellInfo.row['score_colour']
        const sliceLength = 2 * Math.PI * 24
        const sliceOffset = sliceLength * (1 - cellInfo.value / 100)
        const donutChart = (
          '<svg width=60 height=60 style=\"transform: rotate(-90deg)\" focusable=false>' +
            '<circle cx=30 cy=30 r=24 fill=none stroke-width=4 stroke=rgba(0,0,0,0.1)></circle>' +
            '<circle cx=30 cy=30 r=24 fill=none stroke-width=4 stroke=' + sliceColor +
            ' stroke-dasharray=' + sliceLength + ' stroke-dashoffset=' + sliceOffset + '></circle>' +
          '</svg>'
        )
        const label = '<div style=\"position: absolute; top: 50%; left: 50%; ' +
          'transform: translate(-50%, -50%)\">' + cellInfo.row['men_share']+ '</div>'
        return '<div style=\"display: inline-flex; position: relative\">' + donutChart + label + '</div>'
      }"),
                  html = TRUE
                  
                )

),

theme = reactableTheme(
  
  # Vertically center cells
  cellStyle = list(display = "flex", flexDirection = "column", justifyContent = "center"),
  
  headerStyle = list(
    "&:hover[aria-sort]" = list(background = "hsl(0, 0%, 96%)"),
    "&[aria-sort='ascending'], &[aria-sort='descending']" = list(background = "hsl(0, 0%, 96%)"),
    borderColor = "#555"
  )
  
),

highlight = TRUE,
pagination = FALSE

)

```



### How the data was collected

Trade unions are required to complete annual returns. Through these returns, unions provide basic balance sheet summaries and income statements. They also detail their membership breakdown, senior staff, investments, political donations and web-addresses. From the later we can identify each union's twitter handle and thus gather data on their social media presence too. 

So we’ve written a script which does three things:

- Scans the pdfs of union's annual returns, converts them to readable text, and then extracts key information (eg membership numbers and net-assets).
- Scrapes the website address of each union, scans those pages for the unions Twitter handle, and then uses the Twitter API to extract their number of users.
- Analyse the “sentiment” of unions tweets through time, to see if we can detect a shift in industrial relations.

If you’re interested in either of these scripts checkout our GitHub: https://github.com/lobodata/unions
